# Getting and Cleaning Data Project: CodeBook


### Description

The goal of this codebook illustrates how the raw data processed and scripted to produce the tidy data which can be used in study or research projects later. The codebook of this project are not used the files in the Inertial Signals folders. See the README.txt file for the detailed information on the dataset. The files that are used to load data are inclused in the followings:

.test/subject_test.txt  
.test/X_test.txt  
.test/y_test.txt  
.train/subject_train.txt  
.train/X_train.txt  
.train/y_train.txt  

Additional information about the variables, data and transformations can be found in the course project of the Johns Hopkins [Getting and Cleaning Data course.](https://www.coursera.org/course/getdata)

### Data Information

A full description of the data used in this project can be found at [The UCI Machine Learning Repository](http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones ) where the data was obtained.The source data for this project can be found [here.](https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip ) 
The experiments have been carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a smartphone (Samsung Galaxy S II) on the waist. Using its embedded accelerometer and gyroscope, we captured 3-axial linear acceleration and 3-axial angular velocity at a constant rate of 50Hz. The experiments have been video-recorded to label the data manually. The obtained dataset has been randomly partitioned into two sets, where 70% of the volunteers was selected for generating the training data and 30% the test data.
The sensor signals (accelerometer and gyroscope) were pre-processed by applying noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% overlap (128 readings/window). The sensor acceleration signal, which has gravitational and body motion components, was separated using a Butterworth low-pass filter into body acceleration and gravity. The gravitational force is assumed to have only low frequency components, therefore a filter with 0.3 Hz cutoff frequency was used. From each window, a vector of features was obtained by calculating variables from the time and frequency domain.

### The script run_analysis.R performs the 5 sections described in the summary below:

#### Section 1. Merge the training and the test sets to create one data set.

```{r, echo = FALSE, result = "hide"}
library(knitr)
opts_chunk$set(echo = TRUE, cache = TRUE)
```
Clear the workspace
```{r}
rm(list=ls())
```

Set the working directory
```{r}
setwd("C:/Users/giselle/Desktop/Coursera/DATA CLEANING/data/UCI HAR Dataset")
```
After setting the source directory for the files,data sets are loaded:
```{r}
x_train <- read.table("train/X_train.txt")
y_train <- read.table("train/y_train.txt")
subject_train <- read.table("train/subject_train.txt")
```
```{r}
x_test <- read.table("test/X_test.txt")
y_test <- read.table("test/y_test.txt")
subject_test <- read.table("test/subject_test.txt")
```
Concatenate each data set by rows
```{r}
x_data <- rbind(x_train, x_test)
y_data <- rbind(y_train, y_test)
subject_data <- rbind(subject_train, subject_test)
```
Name the variables
```{r}
names(y_data)<- c("activity")
names(subject_data)<-c("subject")
features <- read.table("features.txt")
names(x_data)<- features$V2
```
Combine all data into one data frame
```{r}
dataCombine <- cbind(subject_data, y_data)
Data <- cbind(x_data, dataCombine)
```
```{r}
str(Data,list.len = 20)
```
#### Section 2. Extract only the measurements on the mean and standard deviation for each measurement.

Extract only columns with its names: mean() or std() 
```{r}
subFeatures <- features$V2[grep("mean\\(\\)|std\\(\\)", features$V2)]
selectedNames<-c(as.character(subFeatures), "subject", "activity" )
Data<-subset(Data,select=selectedNames)
```
```{r}
str(Data,list.len = 20)
```
#### Section 3. Use descriptive activity names to name the activities in the data set

Merge data subset with the activityType table to cinlude the descriptive activity names
```{r}
Data$activity <- as.character(Data$activity)
Data$activity[Data$activity == 1] <- "WALKING"
Data$activity[Data$activity == 2] <- "WALKING_UPSTAIRS"
Data$activity[Data$activity == 3] <- "WALKING_DOWNSTAIRS"
Data$activity[Data$activity == 4] <- "SITTING"
Data$activity[Data$activity == 5] <- "STANDING"
Data$activity[Data$activity == 6] <- "LAYING"
Data$activity <- as.factor(Data$activity)
```
```{r}
head(Data$activity,40)
```
#### Section 4. Appropriately label the data set with descriptive activity names.

Use gsub function for pattern replacement to clean up the data labels.
```{r}
names(Data)<-gsub("^t", "time", names(Data))
names(Data)<-gsub("^f", "frequency", names(Data))
names(Data)<-gsub("Acc", "Accelerometer", names(Data))
names(Data)<-gsub("Gyro", "Gyroscope", names(Data))
names(Data)<-gsub("Mag", "Magnitude", names(Data))
names(Data)<-gsub("BodyBody", "Body", names(Data))
```
```{r}
names(Data)
```
#### Section 5. Create a second, independent tidy data set with the average of each variable for each activity and each subject.

Creates a second,independent tidy data set and ouputs it
```{r}
library(plyr)
Data2<-aggregate(. ~subject + activity, Data, mean)
Data2<-Data2[order(Data2$subject,Data2$activity),]
write.table(Data2, file = "tidydata.txt",row.name=FALSE)
```


